Graph Topology Learning from Binary Signals on Graph

### 2022-04-25 poster html copy/poster html.Rmd
Revision in Poster

Data often has  an underlying structure or geometry that can be modeled as a signal on the vertices of a weighted, undirected graph. Traditionally, the emphasis was on using a underlying graph, or network, to understand the properties of signals over the vertices. Recently, there has been a surge in converse problem; of learning a graph structure from a set of signals satisfying some constraints [@xia2021graph] [@dong2016learning] [ @ortega2018graph]. In previous research [@dong2016learning], [@tugnait2021sparse], [@pu2021learning], and [@saboksayr2021accelerated], signals on graph were assumed to follow multivariate gaussian distributions, but there has been little exploration in learning a network from binomial signals. In this work, we develop a noval methodology that allows for learning graph topology given a set of binary signals on the graph. An example of such signals on graph is presented in Figure 1.

@@

\begin{equation}
\label{eq: binaryglm}
\text{logit}(p_{i,j}) = \mathbf{A} h )_i,
\end{equation}

where $\mathbf{A}$ is the adjacency matrix from the graph $G$, $h$ is a vector of latent factors that governs $p_{i, j}$ through $\mathbf{A}$ and assumed to be a standard normal random vector, and $\alpha_j$ is a round specific parameter at stratum $j$, and assumed to be normally distributed with mean of 0, and unknown variance $\sigma^2$.

@@

Consider the proabability mass function for a given $A$ and signal $y = Y[,k]$
\begin{align}
P_{Ah}(y_i) &= P^{y_i}(1-p)^{1-y_i} \\
            &= \left(\frac{e^{A[i,]h^T}}{1 + e^{A[i,]h^T}}\right)\left(1- \frac{e^{A[i,]h^T}}{1 + e^{A[i,]h^T}}\right)^{1-y_i} \\
            &= \left(\frac{e^{A[i,]h^T}}{1 + e^{A[i,]h^T}}\right)\left(\frac{1}{1 + e^{A[i,]h^T}}\right)^{1-y_i} \\
            &= \frac{e^{y(A[i,]h^T)}}{1 + e^{A[i,]h^T}}                     
\end{align}

@@

\begin{align}
\mathcal{L}(h) = \prod_{i=1}^N\frac{e^{y_i(A[i,]h^T)}}{1 + e^{A[i,]h^T}} 
\end{align}

@@

\begin{align}
\log(\mathcal{L}(h)) = \sum_{i=1}^N\left(y_i(A[i,]h^T) - \log(1 + e^{A[i,]h^T})\right)
\end{align}

@@

\begin{equation}
\label{eq:optimization}
\begin{aligned}
&\max_{A,h} \sum_{j = 1}^M\sum_{i=1}^N\left(y_{i,j}(A[i,]h^T) - \log(1 + e^{A[i,]h^T})\right) - \alpha \vert L \vert _ F \\
\textrm{s.t.} \quad  &A_{i,j} = 0 \text{ if } i =j\\
                     &A_{i,j} \geq 0 \text{ if } i \neq j\\
                     &\mathbb{1}h^T = 0 \\
                     &\max(h) \leq a \\
                     &\min(h) \geq b
\end{aligned}
\end{equation}
where $\alpha$ is a tunning parameter for controlling the sparsity and $a$ and $b$ are tunning parameters for restricting the spread of the values of $h$.

@@

\begin{align*}
&1) \textbf{ Input: } \text{ Input a signal } Y.\\
&2) \textbf{ Output: } \text{ Output an estimated } A.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } $i = 1,\ldots, N$\\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } A \textbf{ given } h:\\
&6) \textbf{        Update } h \textbf{ given } A:\\
&7) \textbf{ end for }
\end{align*}
<hr />

@@

In each iteration of step 5) and 6) of Algorithm 1, the optimization program gives a Diciplined Concave Program that can be solved efficiently in Python with the CVXPY libarary. A random Erd\H{o}s-Reyni graph on $N = 20$ nodes and edge probability $p = 0.2$. Heat maps of the ground truth adjacency matrix and estitmated adjacency matrix are given, along with circular embeddings of the ground truth and estimated graphs.

@@

```{r, include=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```
# References

>>>>

Data often has  an underlying structure or geometry that can be modeled as a signal on the vertices of a weighted, undirected graph. Traditionally, the emphasis was on using a underlying graph, or network, to understand the properties of signals over the vertices. Recently, there has been a surge in converse problem; of learning a graph structure from a set of signals satisfying some constraints [@xia2021graph] [@dong2016learning] [ @ortega2018graph]. In previous research [@dong2016learning], [@tugnait2021sparse], [@pu2021learning], and [@saboksayr2021accelerated], signals on graph were assumed to follow multivariate Gaussian distributions, but there has been little exploration in learning a network from binomial signals. In this work, we develop a novel methodology that allows for learning graph topology given a set of binary signals on the graph. An example of such signals on graph is presented in Figure 1.

@@

\begin{equation}
\label{eq: binaryglm}
\text{logit}(p_{i,j}) = (\mathbf{A} h )_i,
\end{equation}

where $\mathbf{A}$ is the adjacency matrix from the graph $G$, $h$ is a vector of latent factors that governs $p_{i, j}$ through $\mathbf{A}$ and assumed to be a standard normal random vector.

@@

Consider the proabability mass function for a given $A$ and signal $y = Y[,k]$
\begin{align}
P_{Ah}(y_i) &= p^{y_i}(1-p)^{1-y_i} \\
            &= \left(\frac{e^{A[i,]h}}{1 + e^{A[i,]h}}\right)^{y_i}\left(1- \frac{e^{A[i,]h}}{1 + e^{A[i,]h}}\right)^{1-y_i} \\
            &= \left(\frac{e^{A[i,]h}}{1 + e^{A[i,]h}}\right)^{y_i}\left(\frac{1}{1 + e^{A[i,]h}}\right)^{1-y_i}   \\
            &= \frac{e^{y_iA[i,]h}}{1 + e^{A[i,]h}}

\end{align}

@@

\begin{align}
\mathcal{L}(h) = \prod_{i=1}^N\left(\frac{e^{y_iA[i,]h}}{1 + e^{A[i,]h}}\right)
\end{align}

@@

\begin{align}
\log(\mathcal{L}(h)) = \sum_{i=1}^N\left(y_i(A[i,]h) - \log(1 + e^{A[i,]h})\right)
\end{align}

@@

\begin{equation}
\label{eq:optimization}
\begin{aligned}
&\max_{A,h} \sum_{j = 1}^M\sum_{i=1}^N\left(y_{i,j}(A[i,]h) - \log(1 + e^{A[i,]h})\right) - \alpha \vert L \vert _ F \\
\textrm{s.t.} \quad  &A_{i,j} = 0 \text{ if } i =j\\
                     &A_{i,j} \geq 0 \text{ if } i \neq j\\
                     &\mathbb{1}h^T = 0 \\
                     &\max(h) \leq a \\
                     &\min(h) \geq b
\end{aligned}
\end{equation}
where $\alpha$ is a tuning parameter for controlling the sparsity and $a$ and $b$ are tuning parameters for restricting the spread of the values of $h$.

@@

\begin{align*}
&1) \textbf{ Input: } \text{ Input a signal } Y.\\
&2) \textbf{ Output: } \text{ Output an estimated } A.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } i = 1,\ldots, N \\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } A \textbf{ given } h:\\
&6) \textbf{        Update } h \textbf{ given } A:\\
&7) \textbf{ end for }
\end{align*}
<hr />

@@

In each iteration of step 5) and 6) of Algorithm 1, the optimization program gives a Disciplined Concave Program that can be solved efficiently in Python with the CVXPY library. A random Erd\H{o}s-Reyni graph on $N = 20$ nodes and edge probability $p = 0.2$. Heat maps of the ground truth adjacency matrix and estimated adjacency matrix are given, along with circular embedding of the ground truth and estimated graphs.

@@

```{r, include=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```
# Acknowledgement
This work is supported by Research, Scholarship, and Creative Activity (RSCA) Award 2021-2022 through Chico State Enterprises.

# References
